---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  cache = TRUE,
  comment = "#>",
  fig.path = "man/figures/",
  out.width = "100%"
)
```

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![R build status](https://github.com/asshah4/card/workflows/R-CMD-check/badge.svg)](https://github.com/asshah4/card/actions)
[![Travis build status](https://travis-ci.com/asshah4/card.svg?branch=master)](https://travis-ci.com/asshah4/card)
[![Github commit frequency](https://img.shields.io/github/commit-activity/w/asshah4/card)](https://github.com/asshah4/card/graphs/commit-activity)
[![Codecov test coverage](https://codecov.io/gh/asshah4/card/branch/master/graph/badge.svg)](https://codecov.io/gh/asshah4/card?branch=master)
<!-- badges: end -->

# __card__

## Cardiovascular and Autonomic Research Design

The inspiration for `card` was to create useful functions and analytical approaches in computational neurocardiology, with a on electrocardiography and epidemiology. 

The areas of focus of this package are the following:

- handling of recurrent and circadian events
- epidemiological focus on effective model building

## Installation

You can install the released version of __card__ from [CRAN](https://CRAN.R-project.org) with:

```{r}
install.packages("card")
```

And the development version from [GitHub](https://github.com/) with:

```{r, eval=FALSE}
# install.packages("devtools")
devtools::install_github("asshah4/card")
```

## Usage

The intent for this `card` is to develop several key functions for the analysis of recurrent or cyclical data, and assistance with epidemiological approaches to regression modeling.

The key features are:

1. Sample datasets
2. Cosinor analysis 
3. Recurrent events
4. Model building

```{r}
library(card)
```

### Datasets

* Cardiovascular disease outcomes with recurrent events
* Vectorcardiography data in a high-risk cohort
* Continuous electrocardiography recordings in a clinical cohort with quantitative stress data

```{r}
# ECG data
data("twins")
head(twins)

# Outcomes data
data("mims")
head(mims)

# VCG data
data("geh")
head(geh)
```

### Cosinor Analysis

This modeling algorithm requires only base `R`, which allows internal flexibility for modeling heuristics and improved efficiency. For the user, it also allows standard modeling tools/approaches, and a flexible user interface that accounts for individual/population analysis and single/multiple component analysis.

There are other cosinor models available in `R`, however they do not allow for simple multiple-component or individual/population analysis in a method that is tidy (for both `tidyverse` and `tidymodels` approaches).

```{r}
m <- cosinor(rDYX ~ hour, twins, tau = c(24, 12))
summary(m)
ggcosinor(m)

head(augment(m))
```

### Recurrent Events

### Model Building