---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  cache = TRUE,
  comment = "#>",
  fig.path = "man/figures/",
  out.width = "100%"
)
```

<!-- badges: start -->
[![Github commit frequency](https://img.shields.io/github/commit-activity/w/shah-in-boots/card)](https://github.com/shah-in-boots/card/graphs/commit-activity)
[![R-CMD-check](https://github.com/shah-in-boots/card/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/shah-in-boots/card/actions/workflows/R-CMD-check.yaml)
[![pkgdown](https://github.com/shah-in-boots/card/actions/workflows/pkgdown.yaml/badge.svg)](https://github.com/shah-in-boots/card/actions/workflows/pkgdown.yaml)
[![CRAN status](https://www.r-pkg.org/badges/version/card)](https://CRAN.R-project.org/package=card)
[![test-coverage](https://github.com/shah-in-boots/card/actions/workflows/test-coverage.yaml/badge.svg)](https://github.com/shah-in-boots/card/actions/workflows/test-coverage.yaml)
[![Codecov test coverage](https://codecov.io/gh/shah-in-boots/card/branch/main/graph/badge.svg)](https://app.codecov.io/gh/shah-in-boots/card?branch=main)
[![](http://cranlogs.r-pkg.org/badges/grand-total/card?color=blue)](https://cran.r-project.org/package=card)
<!-- badges: end -->

# __card__

## Cardiovascular Applications for Research Data

The inspiration for `{card}` was to create useful functions and analytical approaches in computational neurocardiology, with a on electrocardiography and epidemiology. 

The areas of focus of this package are the following:

- handling of recurrent and circadian events
- epidemiological focus on effective model building

## Installation

You can install the released version of `{card}` from [CRAN](https://CRAN.R-project.org) with:

```{r, eval=FALSE}
install.packages("card")
```

And the development version from [GitHub](https://github.com/) with:

```{r, eval=FALSE}
# install.packages("devtools")
devtools::install_github("shah-in-boots/card")
```

## Usage

The intent for this `card` is to develop several key functions for the analysis of recurrent or cyclical data, and assistance with epidemiological approaches to regression modeling.

The key features are:

1. Sample datasets
2. Cosinor analysis 
3. Recurrent events
4. Model building

```{r}
library(card)
```

### Datasets

* Cardiovascular disease outcomes with recurrent events
* Vectorcardiography data in a high-risk cohort
* Continuous electrocardiography recordings in a clinical cohort with quantitative stress data

```{r}
# ECG data
data("twins")
head(twins)

# Outcomes data
data("stress")
head(stress)

# VCG data
data("geh")
head(geh)
```

### Cosinor Analysis

This modeling algorithm requires only base `R`, which allows internal flexibility for modeling heuristics and improved efficiency. For the user, it also allows standard modeling tools/approaches, and a flexible user interface that accounts for individual/population analysis and single/multiple component analysis.

There are other cosinor models available in `R`, however they do not allow for simple multiple-component or individual/population analysis in a method that is tidy (for both `tidyverse` and `tidymodels` approaches).

```{r}
m <- cosinor(rDYX ~ hour, twins, tau = c(24, 12))
summary(m)
ggcosinor(m)

head(augment(m))
```
