---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  cache = TRUE,
  comment = "#>",
  fig.path = "man/figures/",
  out.width = "100%"
)
```

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![R build status](https://github.com/asshah4/card/workflows/R-CMD-check/badge.svg)](https://github.com/asshah4/card/actions)
[![Travis build status](https://travis-ci.com/asshah4/card.svg?branch=master)](https://travis-ci.com/asshah4/card)
[![Github commit frequency](https://img.shields.io/github/commit-activity/w/asshah4/card)](https://github.com/asshah4/card/graphs/commit-activity)
[![Codecov test coverage](https://codecov.io/gh/asshah4/card/branch/master/graph/badge.svg)](https://codecov.io/gh/asshah4/card?branch=master)
<!-- badges: end -->

# __card__

## Cardiovascular and Autonomic Research Design

The goal of __card__ is to create functions and analytical approaches in neurocardiology, with a focus on electrocardiogram data and epidemiological concepts. An important engine within this ecosystem are the circadian analysis tools. Its being developed to fit a _tidy_ approach to statistical analysis.

The areas of focus of this package are the following:

- cardiovascular disease
- circadian rhythms
- electrocardiography (heart rate variability and morphology)
- signal processing

## Installation

Once published, you can install the released version of __card__ from [CRAN](https://CRAN.R-project.org) with:

```{r}
#install.packages("card")
```

And the development version from [GitHub](https://github.com/) with:

```{r}
# install.packages("devtools")
devtools::install_github("asshah4/card")
```

## Usage

This package is intended for analyzing cardiovascular signals and autonomic physiology. The current collection of functions are built practically around this paradigm. The current focus, in `card v0.1.0` is the development of a flexible and tidy analytical approach to circadian rhythms. 

### Cosinor Analysis

Here is an example of the process of fitting a cosinor model to time-series data.

```{r}
library(card)
data(twins)
m <- cosinor(rDYX ~ hour, twins, tau = 24)
summary(m)
```

The data can also be easily plotted for diagnostic purposes.

```{r}
ggcosinor(m)
```

